---
title: "Version Control Assignment"
author: "Abigail Kaye"
date: "`r Sys.Date()`"
output: html_document
---

**dataGenerator.R** 

Set seed for reproducibility
```{r}
set.seed(123)
```

Vector of 5 species names and random selection
```{r}
species_names <- c("Kaye", "Mya", "Lauren", "Susie", "King")
species <- sample(species_names, 100, replace = TRUE)
```

Vector of 100 Limb_width values (using log-normal distribution)
```{r}
limb_width <- rlnorm(100, meanlog = 1, sdlog = 0.5)
```

Vector of 100 Limb_length values (using normal distribution)
```{r}
limb_length <- rnorm(100, mean = 100, sd = 20)
```

Vector of 3 observer names and random selection
```{r}
observer_names <- c("John Smith", "Emily Johnson", "Michael Lee")
observer <- sample(observer_names, 100, replace = TRUE)
```

Combine all vectors into a data frame
```{r}
data <- data.frame(
  Organism = species,
  Limb_width = round(limb_width, 2),
  Limb_length = round(limb_length, 2),
  Observer = observer
)

```

Display the first few rows of the data
```{r}
head(data)
```

Export the data frame to a CSV file
```{r}
write.csv(data, "measurements.csv", row.names = FALSE)
```

**volumeEstimato.R**

```{r}
library(dplyr)
library(readr)
```

To calculate volume and add it as a new column I use a simplified cylinder volume formula: V = π * r^2 * h
r is half the limb width, and h is the limb length
```{r}
data_with_volume <- data %>%
  mutate(Volume = pi * (Limb_width/2)^2 * Limb_length)
```

Round the Volume to 2 decimal places for readability
```{r}
data_with_volume <- data_with_volume %>%
  mutate(Volume = round(Volume, 2))
```

Write the updated data back to measurements.csv, overwriting the original file
```{r}
write_csv(data_with_volume, "measurements.csv")
```

Display the first few rows of the updated data
```{r}
print(head(data_with_volume))
```

Required Libararies
```{r}
library(dplyr)
library(ggplot2)
```

**1. Load the revised measurements.csv**
```{r}
data <- read.csv("measurements.csv")
```

**2.Sort the data by species, observer, and limb volume**
```{r}
sorted_data <- data %>%
  arrange(Organism, Observer, Volume)
```

**3.Create a table showing the average volume for each species**
```{r}
avg_volume_by_species <- data %>%
  group_by(Organism) %>%
  summarise(avg_volume = mean(Volume))
```

**4. Create a table showing the number of observations for each combination of species and observer**
```{r}
obs_count <- data %>%
  group_by(Organism, Observer) %>%
  summarise(count = n())
```

**5. Create a box plot to compare key features of the distributions of volumes for each species**
```{r}
ggplot(data, aes(x = Organism, y = Volume)) +
  geom_boxplot() +
  labs(title = "Distribution of Limb Volumes by Species",
       x = "Species",
       y = "Volume")
```

**6. Create a multi-panel plot showing frequency histograms showing the distribution of limb volume for each species. Note that this may require you to add code to you R markdown file to change the data from the wide to long format**
```{r}
ggplot(data, aes(x = Volume)) +
  geom_histogram(bins = 30) +
  facet_wrap(~Organism, scales = "free_y") +
  labs(title = "Distribution of Limb Volumes by Species",
       x = "Volume",
       y = "Frequency")
```





